---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

knitr::opts_chunk$set(echo=TRUE)

## Loading and preprocessing the data

```{r}
data = read.csv("activity.csv")
str(data)
summary(data)

```


## What is mean total number of steps taken per day?
```{r total steps}
sum <- tapply(data$steps, data$date, sum, na.rm=TRUE)
```
Total number of steps per day by day are `r sum`

```{r histogram}
hist(sum, main = "Histogram of the total number of steps taken each day", xlab = "Total No. of Steps", ylab = "Frequency", col = "red")
```

```{r mean and median}
steps_mean = mean(sum)
steps_median = median(sum)
```
Mean number of the total number of steps taken per day is `r steps_mean`.
Median number of the total number of steps taken per day is `r steps_median`. 

## What is the average daily activity pattern?
```{r activity pattern}
daily_avg = tapply(data$steps, data$interval, mean, na.rm=TRUE)
interval_vector = as.vector(seq(1:288))
plot(interval_vector, daily_avg, type="l", xlab="Interval", ylab="Daily Average", main="5-minute interval vs. the average number of steps taken")
```

```{r max steps}
max_interval =  which.max(daily_avg)
```
Daily interval containing the maximum number of steps is `r max_interval`

## Imputing missing values
```{r missing NAs}
missing = sum(is.na(data$steps))
```
The total number of missing values in the data set are `r missing`

```{r replace missing NAs}
newdata = data # make a copy of the data
newdata$steps[is.na(newdata$steps)] = mean(daily_avg)
newdaily_avg = tapply(newdata$steps, newdata$interval, mean)
summary(newdata)
```

```{r new total steps}
newsum <- tapply(newdata$steps, newdata$date, sum, na.rm=TRUE)
```
Total number of steps per day by day are `r newsum`

```{r newhistogram}
hist(newsum, main = "Histogram of the total number of steps taken each day", xlab = "Total No. of Steps", ylab = "Frequency", col = "red")
```

```{r new mean and median}
newsteps_mean = mean(newsum)
newsteps_median = median(newsum)
```
Mean number of the total number of steps taken per day is `r newsteps_mean`.
Median number of the total number of steps taken per day is `r newsteps_median`. 

Difference between New mean and old mean is `r steps_mean - newsteps_mean`.
Difference between New median and old median is `r steps_median - newsteps_median`.
Yes, this is a difference, but it depends on how the NAs are imputed. Imputing with medians produced a more normal distribution.

## Are there differences in activity patterns between weekdays and weekends?
```{r weekday and weekend}
newdata1 = newdata
newdata1$date = as.Date(newdata1$date)
weekdays <- c('Monday', 'Tuesday', 'Wednesday', 'Thursday', 'Friday')
newdata1$whichday = factor((weekdays(newdata1$date) %in% weekdays), level=c(FALSE, TRUE), labels=c('weekend', 'weekday'))
str(newdata1)
```

```{r panel plot}
weekday_data = subset(newdata1, whichday=="weekday")
weekend_data = subset(newdata1, whichday=="weekend")
weekday_avg = tapply(weekday_data$steps, weekday_data$interval, mean)
weekend_avg = tapply(weekend_data$steps, weekend_data$interval, mean)
par(mfrow=c(2,1))
plot(interval_vector, weekday_avg, type="l", main="weekday plot", xlab = "Interval", ylab="No of Steps")
plot(interval_vector, weekend_avg, type="l", main="weekend plot", xlab = "Interval", ylab="No of Steps")
#plot(newdata1$steps[newdata1$whichday=="weekend"])
```
